---
title: "Brexit"
author: "Rebekah Chuang"
output: html_document
---

```{r load-packages, message = FALSE, echo = FALSE}
library(tidyverse)
library(scales)
```

In September 2019, YouGov survey asked 1,639 GB adults the following question:

> In hindsight, do you think Britain was right/wrong to vote to leave EU?
>
> -   Right to leave  
> -   Wrong to leave  
> -   Don't know

The data from the survey is in `data/brexit.csv`.

```{r message = FALSE}
brexit <- read_csv("data/brexit.csv")
brexit
```

In the course, we made the following visualisation.

```{r original-plot}
brexit <- brexit %>%
  mutate(
    region = fct_relevel(region, "london", "rest_of_south", "midlands_wales", "north", "scot"),
    region = fct_recode(region, London = "london", `Rest of South` = "rest_of_south", `Midlands / Wales` = "midlands_wales", North = "north", Scotland = "scot")
  )

ggplot(brexit, aes(y = opinion, fill = opinion)) +
  geom_bar() +
  facet_wrap(~region, nrow = 1, labeller = label_wrap_gen(width = 12)) +
  guides(fill = FALSE) +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = NULL, y = NULL
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  theme_minimal()
```

In this application exercise we tell different stories with the same data.

### Exercise 1 - Free scales

Add `scales = "free_x"` as an argument to the `facet_wrap()` function.
How does the visualisation change?
How is the story this visualisation telling different than the story the original plot tells?

```{r set-scales-to-free-x}
ggplot(brexit, aes(y = opinion, fill = opinion)) +
  geom_bar() +
  facet_wrap(~region,
    nrow = 1, labeller = label_wrap_gen(width = 12),
    scales = "free_x"
  ) +
  guides(fill = FALSE) +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = NULL, y = NULL
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  theme_minimal()
```

After adding `scales = "free_x"` as an argument to the `facet_wrap()` function, the x-axis scales in each facet vary independently based on the data. This adjustment allows users to interpret the data within each facet more effectively and gain a better understanding of the proportions of each answer, as well as how the results compare across different areas.

In the original plot, the bars for **London** and **Scotland** appeared relatively small compared to other areas. However, when looking at the updated plots, it becomes evident that the proportion of answers categorized as **Wrong** is nearly the same for each area.

### Exercise 2 - Comparing proportions across facets

First, calculate the proportion of wrong, right, and don't know answers in each category and then plot these proportions (rather than the counts) and then improve axis labeling.
How is the story this visualisation telling different than the story the original plot tells?
**Hint:** You'll need the **scales** package to improve axis labeling, which means you'll need to load it on top of the document as well.

```{r calculate-proportion}
brexit_prob = brexit %>% 
  count(region, opinion) %>%
  group_by(region) %>% 
  mutate(prob = n/sum(n))
brexit_prob
```
```{r proportion-facet}
ggplot(brexit_prob, aes(x = prob, y = opinion, fill = opinion)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~region,
             nrow = 1) +
  guides(fill = FALSE) +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = NULL, y = NULL
  ) +
    scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  theme_minimal()
```

### Exercise 3 - Comparing proportions across bars

Recreate the same visualisation from the previous exercise, this time dodging the bars for opinion proportions for each region, rather than faceting by region and then improve the legend.
How is the story this visualisation telling different than the story the previous plot tells?

```{r dodged-bar-plot}
ggplot(brexit_prob, aes(x = prob, y = region, fill = opinion)) + 
  geom_bar(stat = "identity",
           position = "dodge") +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = "Proportion", y = "Region",
    fill = "Opinion"
  ) +
    scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  ), breaks = c("Wrong", "Right", "Don't know")) +
  theme_minimal()
```

If we use dodging bars instead of faceting by region, users can better compare the proportions between each region. For example, we can clearly see that **Scotland** has the highest proportion of answers marked as `Wrong` compared to other regions. The proportion of `Right` answers is similar between North and the **Rest of the South**.
